Circular Distributional Trees and Forests with an Application to Wind Direction Forecasting

### 2019-11-20 pkg/circtree/inst/wind_paper/draft/draft.Rnw
Include 'new draft' for abstract
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2887 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: b68ab7b74a4372cd65aa4a31e380ba115b497d33

\Abstract{
As circular data can be found in many different subject areas accounting for periodicity has 
become an important aspect in establishing appropriate statistical models.
For probabilistic modeling of circular data the von Mises distribution is widely used.
To capture how its parameters change with covariates, a distributional regression tree model
is proposed as an alternative to more commonly-used additive models. The resulting
distributional trees are easy to interpret, can detect non-additive effects, and select
covariates and their interactions automatically. In order to stabilize and regularize
the model an ensemble of distributional trees is combined yielding a distributional
forest which allows for modeling smooth effects as well.
For illustration, hourly wind direction forecasts are obtained at two airports in Austria
(Vienna and Innsbruck) based on a set of meteorological measurements from each of the 
airports and surrounding weather stations.
}
\Keywords{Distributional Trees and Forests; Circular Response; Von Mises Distribution}

@@

\begin{figure}[t]
%\centering
\hspace{-1cm}
\begin{minipage}{.5\textwidth}
  %\centering
  \includegraphics[width=1.2\linewidth]{density_linear.pdf}
  \end{minipage}%
\begin{minipage}{.5\textwidth}
  %\centering
  \includegraphics[width=1.2\linewidth]{density_circular.pdf}
\end{minipage}
\caption{Illustration of a distributional model fitted to circular data by 
  maximum likelihood-estimation employing the von Mises distribution. 
  Left panel: linear illustration. Right panel: circular illustration.
  In both panels the empirical histogram (gray) and fitted density (red line) are depicted along 
  with the estimated location parameter (red hand).
  }
  \label{fig:densities} 
  %\includegraphics[width=4cm]{density_vM.pdf}
\end{figure}
\fixme{update figure 2d to 3d, using data of example tree}

>>>>

\Abstract{
As circular data can be found in many different areas, the development of
appropriate statistical models accounting, i.a., for periodicity has become a
challenging and important research field. For probabilistic forecasting
typically the von Mises distribution is employed in an additive model
framework. Combining their underlying distributional setup with tree-based
estimation methods, circular distributional trees will be presented. The resulting
distributional trees are easy to interpret, can detect non-additive effects,
and select covariates and their interactions automatically. By combining
ensembles of circular trees a stabilized circular forest can build, which
automatically regularizes the estimation and, hence, allows for modeling smooth effects.

To validate the proposed circular trees and forests, a case study on
probabilistic forecasting of wind direction will be shown for two airports in
Austria. In addition to the tree-based models, alternative commonly used
approaches will be employed to compare their predictive performances to the
proposed circular trees and forest.

%As circular data can be found in many different subject areas accounting for periodicity has 
%become an important aspect in establishing appropriate statistical models.
%For probabilistic modeling of circular data the von Mises distribution is widely used.
%To capture how its parameters change with covariates, a distributional regression tree model
%is proposed as an alternative to more commonly-used additive models. The resulting
%distributional trees are easy to interpret, can detect non-additive effects, and select
%covariates and their interactions automatically. In order to stabilize and regularize
%the model an ensemble of distributional trees is combined yielding a distributional
%forest which allows for modeling smooth effects as well.
%For illustration, hourly wind direction forecasts are obtained at two airports in Austria
%(Vienna and Innsbruck) based on a set of meteorological measurements from each of the 
%airports and surrounding weather stations.
}
\Keywords{Distributional Trees and Forests; Circular Response; von Mises Distribution}

@@

\begin{figure}[t]
%\centering
%\hspace{-1cm}
\begin{minipage}{.05\textwidth}
\hfill
\end{minipage}%
\begin{minipage}{.45\textwidth}
  %\centering
  \vspace{-1em}
  \includegraphics[width=1.05\linewidth]{density_linear.pdf}
  \end{minipage}%
\begin{minipage}{.45\textwidth}
  %\centering
  \includegraphics[width=1.2\linewidth]{density_circular.pdf}
\end{minipage}
\begin{minipage}{.05\textwidth}
\hfill
\end{minipage}%
\caption{Illustration of a distributional model fitted to circular data by 
  maximum likelihood-estimation employing the von Mises distribution. 
  Left panel: linear illustration. Right panel: circular illustration.
  In both panels the empirical histogram (gray) and fitted density (red line) are depicted along 
  with the estimated location parameter (red hand).
  }
  \label{fig:densities} 
  %\includegraphics[width=4cm]{density_vM.pdf}
\end{figure}
\fixme{update figure 2d to 3d, using data of example tree}
